---
title: "analysis"
author: "Maryam Sanni"
format: html
editor: visual
---

The dataset have been cleaned and is ready to be used for analysis

There are 15 datasets here. These are:

- dailyActivity_merged: This is a summary of all activity measurements in a day

- dailyintensities_merged: intensity of activities(time and distance)

- dailySteps_merged: Amount of steps taken each day

- heartrate_seconds_merged: heart rate per seconds

- hourlyCalories_merged: Calories burnt each hour

- hourlyIntensities_merged: Average and total intensities per hour

- hourlySteps_merged: Total steps taken per hour

- minuteCaloriesNarrow_merged: Amount of calories burnt per minute

- minuteMETsNarrow_merged: The MET value per minute

- minuteSleep_merged: Sleep intensity per minute

- minuteStepsNarrow_merged: Steps taken per minute

- sleepDay_merged: number of time slept, total minutes slept and total time in bed

- weightLogInfo_merged: weight(kg, pounds), BMI and if it was manually reported

## Lets get some insights!

**What day do people take the most steps**

create a new column with each day of the week
Sunday-1, Monday-2, Tuesday-3, Wednesday-4, Thursday-5, Friday-6, Saturday-7

```{r}
dailyActivity_merged$day <- wday(dailyActivity_merged$ActivityDate)
```


```{r}
library(tidyr)
highest_step <- dailyActivity_merged %>% 
  group_by(day) %>% 
  summarise(Total_Steps = sum(TotalSteps), averageSteps = mean(TotalSteps), TotalCalories = sum(Calories), averageCalories = mean(Calories)) %>% 
  select(day, Total_Steps, averageSteps, TotalCalories, averageCalories) %>% 
  mutate(Total_Steps = Total_Steps*0.001)

print( highest_step, n = nrow(highest_step))
```

```{r}

days_of_week <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")

highest_step$day <- factor(highest_step$day,
                           levels = unique(highest_step$day),
                           labels = days_of_week
)

highest_step <- gather(highest_step, key = variable, value = "value", -day)

# create bar chart
ggplot(highest_step, aes(x = day, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "steps taken (in thousands)", x = "Day", y = "Total steps (seconds)") +
 # scale_fill_manual(values = c("Sunday" = "red", "Monday" = "yellow", "Tuesday" = "blue", "Wednesday" = "purple", "Thursday" = "brown", "Friday" = "green", "Saturday" = "pink"), labels = days_of_week) +
  theme_minimal()
```
From the table below, people who use the app take the most steps on Tuesdays but the least steps on sundays. 

**What day do people take the most steps**

```{r}

```





